---
title: Interactive mapping with RShiny
author: ''
date: '2022-07-20'
slug: []
categories: []
tags: 
  - shiny
  - leaflet
  - tmap
  - sf
Description: ''
Tags: []
output:
  blogdown::html_page:
    toc: true
Categories: []
DisableComments: no
---

# A journey in maps and apps...
<br>
1. The basics of programming for interactivity: *Reactivity*
<br>
2. R package options for mapping interactively in R, with pros and cons
<br>
3. Make an interactive map and share in a web app
<br>

**Note**: This journey is thanks to the awesome book [Mastering Shiny](https://mastering-shiny.org/index.html)
<br>

## Why interactive and why Shiny?

**Open Science**

  - Interactive web apps are a great way to make our research **open**, **transparent**, and **engaging**
  - Move beyond static papers

**Shiny**
  
  - Performative apps without HTML, javascript, CSS...
  - Range of users - academics to tech companies 
  - Need to be hosted on a web server

**Server options**

  - [shinyapps.io](https://www.shinyapps.io/): free or subscription options
  - [See here for more info on Rshiny servers](https://shiny.rstudio.com/articles/shiny-server.html)
  - There are other 'non-R' server options out there too

## App fundamentals

Interactivity requires *Reactivity*

What is **reactive programming**?

  - Outputs *react* to inputs

    - Means computations are only done when necessary (moving beyond **functions** to **reactions**)
    - Reactive contexts are **lazy** and **cached**


  - It's different from how we normally code our scripts:
  
    - Commands vs. recipes
    - Imperative vs. declarative
    - Assertive vs. passive-aggressive

## Shiny

{shiny} provides a framework for **reactive programming**:


  - User-interface (ui): how it looks. Takes *inputs* and displays *outputs*
  
  - Server: has the recipe to turn *inputs* into *outputs*

*hint*: if you have {shiny} installed, just type `shinyapp` in your R script to insert boilerplate shiny app code and get started quickly

```{r, eval=FALSE}
library(shiny)
ui <- fluidPage(
  
)
server <- function(input, output, session) {
  
}
shinyApp(ui, server)
```

## A simple histogram

```{r}
hist(rnorm(25))
```


## A reactive histogram

```{r, eval=FALSE}
library(shiny)
ui <- fluidPage(
  numericInput(inputId = "n", "Sample size", value = 25), 
  plotOutput(outputId = "hist")
)
server <- function(input, output, session) {
  output$hist <- renderPlot({
    hist(rnorm(input$n))
    })
}
shinyApp(ui, server)
```

**Note** There's an additional reactive tool, *reactive expressions*. *Reactive expressions* can be used to eliminate redundant code in your app, thereby improving efficiency. Find more information [here](https://mastering-shiny.org/reactivity-intro.html).

## But what about maps?

My go-to for mapping in R these days is {tmap}

One of the reasons is that I can make either interactive or static, publication-quality maps with ease. Just set the 'mode'.

```{r, warning = F}
library(tmap)
data('World')
tmap_mode('view')
qtm(World, 'name') # qtm stands for 'quick thematic mapper'
```

## Share with shiny

If we're in a hurry, we can put our interactive {tmap} in a {shiny} app to share

```{R, eval=FALSE}
library(shiny)
library(tmap)
data('World')
ui <- fluidPage(
  tmapOutput('map')
)
server <- function(input, output, session) {
  output$map <- renderTmap({
    qtm(World, 'name')})
}
shinyApp(ui, server)
```

## Add some reactivity

Let's add a widget to allow users to choose their country from a list

```{R, eval=FALSE}
library(shiny)
library(tmap)
library(dplyr)
data('World')
ui <- fluidPage(
  tmapOutput('map'),
  selectInput('var', 'Select country',
              choices = c('Global', as.character(World$name)),
              selected = 'Global')
)
server <- function(input, output, session) {
  output$map <- renderTmap({
    if(input$var == 'Global'){
      qtm(World, 'name')
    }else{
      qtm(filter(World, name == input$var))
    }
  })
}
shinyApp(ui, server)
```


## Customising and scaling your interactive map

If you want to **customise** your map with more *features*, the [leaflet](https://rstudio.github.io/leaflet/) R package offers some nice functionality (there are others too)

{tmap} and {leaflet} both use **'scalable vector graphics'** (SVG) to visualise maps on the web

  - This is fine for simple apps, but if you need to render large datasets or do complex computations they will be really slow

**Scaling-up** your map app might require graphics to be rendered in **'WebGL'** instead of SVG


  - There are several R packages to try including [mapdeck](https://symbolixau.github.io/mapdeck/articles/mapdeck.html) and [rdeck](https://github.com/anthonynorth/rdeck)
  - These usually require API tokens to access basemaps from [mapbox](https://www.mapbox.com/)

Should I go **SVG** or **WebGL**?

  - Generally I've found there is a **feature vs. speed trade-off**
  - Recommend start with **SVG** and if too slow then move to **WebGL** 

## Tips and tricks

- Lots of complex polygons? Use [rmapshaper](https://cran.r-project.org/web/packages/rmapshaper/vignettes/rmapshaper.html) to simplify, this can really **speed up** a slow map app
- App still slow? Use [profvis](https://rstudio.github.io/profvis/) to profile your script and find the bottleneck
- Have you reached the **'cliff of complexity'**?
Check out [Engineering production grade shiny apps](https://engineering-shiny.org/successful-shiny-app.html)


## Leaflet example

Follow the R script 'leaflet-map.R' to build an interactive {leaflet} map

Next follow the R script 'leaflet-app.R' to turn the interactive {leaflet} map into a share-able and dynamic {shiny} web app


