---
title: Creating a Geospatial blog with blogdown Part 3. A Raster map.
author: UQGSAC
date: '2021-10-28'
slug: a-raster-map
categories:
  - spatial
  - visualisation
series:
  - series-setup  
tags:
  - RStats
  - raster
---

An example raster visualisation.

## Load the packages

```{r}
library(terra)
```

## Import the data

The data comes from the Bureau of Meteorology website, it is a raster file of average annual rainfall. We've put the file into a `data` directory, inside the blog post's directory.

```{r}
rain <- rast("data/rainan.txt")
```

## Inspect

```{r}
rain
```

One single band, by default with the WGS 84 CRS.

The average rainfall for the whole raster is `r rain |> values() |> mean() |> round(2)` mm.

## Visualise

Make sure to add a caption to visualisations, and some alternative text if needed!

```{r}
#| fig.cap = "Average annual rainfall in mm (1980 to 2010)",
#| fig.alt = "Map of Australia and surrounds' average annual rainfall, ranging from 0 
#|  to about 4300 mm."
plot(rain)
```